---
title: "ABIA"
output: html_document
---

```{r echo=FALSE}
abia = read.csv('C:\\Users\\cherr\\STA380\\data\\ABIA.csv',header = TRUE)
library(ggplot2)
```

## Trying to identifyt he biggest carrier in Austin

```{r echo=FALSE}
ggplot(abia, aes(UniqueCarrier))+
  geom_bar(stat="count", width=0.7, fill="steelblue") +
  labs(title= "Biggest carrier in Austin", 
     x="Carrier", y = "Number of flights")
```

### Looks like WN and AA are the two biggest carriers in 2008

## Number of Flights arriving in Austin

```{r echo=FALSE}
aus_arrival = abia[abia$Dest == 'AUS',]
y <-aggregate(cbind(count = FlightNum) ~ Origin, 
          data = aus_arrival, 
          FUN = function(x){NROW(x)})
y_sorted <- y[order(-y$count),][1:10,]

ggplot(y_sorted, aes(x = Origin, y = count))+
  geom_bar(stat = 'identity',width=0.7,fill = 'tomato3')+
  labs(title= "Top 10 origins of flights to Austin", 
       x="Origin", y = "Number of flights")
```

### These are the top 10 origins with flights landing in Austin. DAL and DFW (Dallas) hosts majority of flights

## Flights departing from Austin

```{r echo=FALSE}
aus_dest = abia[abia$Origin == 'AUS',]
x <-aggregate(cbind(count = FlightNum) ~ Dest, 
              data = aus_dest, 
              FUN = function(x){NROW(x)})
x_sorted <- x[order(-x$count),][1:10,]


ggplot(x_sorted, aes(x = Dest, y = count))+
  geom_bar(stat = 'identity',width=0.7,fill = 'yellow')+
  labs(title= "Top 10 destination of flights from Austin", 
       x="Destinations", y = "Number of flights")
```

### Looks like most of the flights starting from Austin are directed towards the above mentioned top 10 destinations. Again majority of the flights are for Dallas

## Now lets see the frequency of the flights from these top 10 locations to Austin

```{r echo=FALSE}
top_10 = aus_arrival[aus_arrival$Origin %in% c('ATL','DAL','DEN','DFW','HOU','IAH','JFK','LAX','ORD','PHX'),]
top_10_x <-aggregate(cbind(count = FlightNum) ~ Origin + Month, 
              data = top_10, 
              FUN = function(x){NROW(x)})

ggplot(top_10_x, aes(x=factor(Month), y=count, group=Origin)) +
  geom_line(aes(color=Origin))+
  geom_point(aes(color=Origin))+
  scale_color_brewer(palette="Dark2")+
  labs(title= "Number of flights for top 10 origins throughout the year", 
       x="Month", y = "Number of flights")
```

### Looks like number of flights originating from Dallas are reducing from mid of the year

```{r echo=FALSE}
top_10_dest = aus_dest[aus_dest$Dest %in% c('ATL','DAL','DEN','DFW','HOU','IAH','JFK','LAX','ORD','PHX'),]
top_10_y <-aggregate(cbind(count = FlightNum) ~ Dest + Month, 
                     data = top_10_dest, 
                     FUN = function(x){NROW(x)})

ggplot(top_10_y, aes(x=factor(Month), y=count, group=Dest)) +
  geom_line(aes(color=Dest))+
  geom_point(aes(color=Dest))+
  scale_color_brewer(palette="Dark2")+
  labs(title= "Number of flights for top 10 destinations throughout the year", 
       x="Month", y = "Number of flights")
```
### We can see a similar drop in Austin to Dallas flights from mid of the year. Looks like something went wrong here

## Lets see why there was a sudden drop in flights originating and departing from Dallas to Austin

```{r echo=FALSE}
dallas = abia[(abia$Origin == 'DAL' | abia$Dest == 'DAL'),]
dallas_carrier = aggregate(cbind(count = FlightNum) ~ UniqueCarrier + Month, 
                           data = dallas, 
                           FUN = function(x){NROW(x)})

ggplot(dallas_carrier, aes(x=factor(Month), y=count, group=UniqueCarrier)) +
  geom_line(aes(color=UniqueCarrier))+
  geom_point(aes(color=UniqueCarrier))+
  scale_color_brewer(palette="Dark2")+
  labs(title= "Number of flights originating and landing in dallas by carrier", 
       x="Month", y = "Number of flights")
```
### MQ carrier stopped its service after June and this clearly explains the drop in the Dallas flights

## Checking frequency of short and long distance flights
### Flights originating from Austin
```{r echo=FALSE}
abia$type <- ifelse(abia$Distance < 701, 'Short distance','Long distance')
dist <- abia[abia$Origin == 'AUS',]

ggplot(data=dist, aes(x=factor(Month), fill=type)) +
  geom_bar(stat="count", position=position_dodge())+
  scale_fill_brewer(palette="Paired")+
  labs(title= "Number of short/long distance flights originating from Austin", 
      x="Month", y = "Number of flights")
```

### Flights landing in Austin
```{r echo=FALSE}
dist1 <- abia[abia$Dest == 'AUS',]

ggplot(data=dist1, aes(x=factor(Month), fill=type)) +
  geom_bar(stat="count", position=position_dodge())+
  scale_fill_brewer(palette="Paired")+
  labs(title= "Number of short/long distance flights landing in Austin", 
       x="Month", y = "Number of flights")
```

## Taxiout time of flights starting from Austin to the top 10 airports
```{r echo=FALSE}
ggplot(top_10_dest, aes(factor(DayOfWeek), TaxiOut))+
  geom_boxplot(varwidth=T, fill="plum") +
  labs(title= "TaxiOut time by Day of the week", 
       x="Day of the week", y = "TaxiOut time")
```
### Looks like day of week really doesn't matter here

```{r echo=FALSE}
ggplot(top_10_dest, aes(UniqueCarrier, TaxiOut))+
  geom_boxplot(varwidth=T, fill="orange") +
  labs(title= "TaxiOut time by Carrier", 
       x="Carrier", y = "TaxiOut time")
```

### Bigger airlines like AA and WN are given higher preference to take off and hence have lower taxiOut time

```{r echo=FALSE}
ggplot(top_10_dest, aes(factor(round(DepTime/100, digits =0)), TaxiOut))+
  geom_boxplot(varwidth=T, fill="orange") +
  labs(title= "TaxiOut time by Actual dept time", 
       x="actual dept time", y = "TaxiOut time")
```

### Looks like morning 6-8 and afternoon 3-6 is the busisest time

## Reasons of delay for the flights originating from Austin - to top 10 airports

### %flights with Carrier Delay by carrier

```{r echo=FALSE}
carr_delay1 <- aggregate(FlightNum ~ UniqueCarrier,data=top_10_dest,function(x){NROW(x)})
carr_delay2 <- aggregate(CarrierDelay ~ UniqueCarrier,data=top_10_dest,function(x){NROW(is.na(x))})
carr_delay3 <- cbind(carr_delay1,carr_delay2)
carr_delay3$CarrierDelay_percent <- (carr_delay3$CarrierDelay/carr_delay3$FlightNum)*100
carr_delay3 <- carr_delay3[-3]

ggplot(carr_delay3, aes(x=UniqueCarrier, y=CarrierDelay_percent)) +
  geom_bar(stat = 'identity',width=0.7,fill = 'tomato3')+
  labs(title= "Flights with carrier delay by carrier", 
       x="Carrier", y = "% flights with carrier delay")
```

### % flights with weather delay by month

```{r echo=FALSE}
weather_delay1 <- aggregate(FlightNum ~ Month,data=top_10_dest,function(x){NROW(x)})
weather_delay2 <- aggregate(WeatherDelay ~ Month,data=top_10_dest,function(x){NROW(is.na(x))})
weather_delay3 <- cbind(weather_delay1,weather_delay2)
weather_delay3$WeatherDelay_percent <- (weather_delay3$WeatherDelay/weather_delay3$FlightNum)*100
weather_delay3 <- weather_delay3[-3]

ggplot(weather_delay3, aes(x=factor(Month), y=WeatherDelay_percent)) +
  geom_bar(stat = 'identity',width=0.7,fill = 'steelblue')+
  labs(title= "Flights with weather delay by month", 
       x="Month", y = "% flights with weather delay")
```